---
title: "PhyloTags"
author: "Laura Salazar-Jaramillo"
date: "2023-01-19"
output:
  html_notebook:
    fig_caption: yes
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

## Preprocessing

A phyloseq object is first created for each marker from the count tables, trees, taxonomy and sequences

```{r cache=TRUE,results='hide',warning=FALSE,echo=FALSE,message=FALSE,cache=TRUE}
#setwd("G:/BackUp_vidarium/BackUp_2023enero/proyectos/phylotags/")
source("G:/BackUp_vidarium/BackUp_2023enero/proyectos/phylotags/articulo/scripts/phylotags_preprocessing.R")
source("G:/BackUp_vidarium/BackUp_2023enero/proyectos/phylotags/articulo/scripts/phylotags_diversity.R")
#source("G:/BackUp_vidarium/BackUp_2023enero/proyectos/phylotags/articulo/scripts/phylotags_plots.R")
#library(kableExtra)
library(dplyr)
library(phyloseq)

```

## Diversity Analyisis

Two models were run, a binary model and a continuous model for each genus of the selected families within Clostridia

```{r,diversity,results='hide',warning=FALSE,echo=FALSE,message=FALSE,cache=TRUE}
#source("G:/BackUp_vidarium/BackUp_2023enero/proyectos/phylotags/articulo/scripts/phylotags_diversity.R")

```

## Binary Model for Families

### dnaK

```{r, dnak_bin_fam,results='show',warning=FALSE,echo=FALSE,message=FALSE}

dnak_tax_classfam <- as.data.frame(unique(phyloseq::tax_table(dnak_physeq_dada2)[,c("class","family")]))
dnak_tax_classfam$family<- str_replace_all(dnak_tax_classfam$family, "-","_")
dnak_glmbin_fam.sigtab %>%
    left_join(dnak_tax_classfam,by=c("taxa"="family")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia")) %>%
    mutate(taxa=str_replace_all(taxa, "f__",""))

```

### gyrB

```{r, gyrB_bin_fam,results='show',warning=FALSE,echo=FALSE,message=FALSE}
gyrB_tax_classfam <- as.data.frame(unique(phyloseq::tax_table(gyrB_physeq_dada2)[,c("class","family")]))
gyrB_tax_classfam$family<- str_replace_all(gyrB_tax_classfam$family, "-","_")
gyrB_glmbin_fam.sigtab %>%
    left_join(gyrB_tax_classfam,by=c("taxa"="family")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))

```


```

## Continuous Model for Families

### dnaK

```{r, dnak_cont_fam_table,results='show',warning=FALSE,echo=FALSE,message=FALSE}

dnak_glmcont_fam.sigtab %>%
    left_join(dnak_tax_classfam,by=c("taxa"="family")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia")) %>%
    mutate(taxa=str_replace_all(taxa, "f__","")) 

```

## Significant families for dnaK with corresponding plot for 16S

```{r, dnak_cont_fam_plot,results='show',warning=FALSE,echo=FALSE,message=FALSE}

oscillo_fam.plot

```

### gyrB

```{r, gyrB_cont_fam,results='show',warning=FALSE,echo=FALSE,message=FALSE}

gyrB_glmcont_fam.sigtab %>%
    left_join(gyrB_tax_classfam,by=c("taxa"="family")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
    mutate(taxa=str_replace_all(taxa, "f__",""))

```

## Significant families for gyrB with corresponding plot for 16S

```{r, gyrB_cont_fam_plot,results='show',warning=FALSE,echo=FALSE,message=FALSE}

gyrB_fam.plot

```

### 16S

```{r, 16S_cont_fam,results='show',warning=FALSE,echo=FALSE,message=FALSE}

ssu_glmcont_fam.sigtab %>%
    left_join(ssu_tax_classfam,by=c("taxa"="family")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia")) 

```

## Binary Model for Genera

### dnaK

```{r, dnak_bin_gen,results='show',warning=FALSE,echo=FALSE,message=FALSE}
dnak_tax_famgenus <- as.data.frame(unique(phyloseq::tax_table(dnak_physeq_dada2)[,c("class","family","genus")]))
dnak_tax_famgenus$genus<- str_replace_all(dnak_tax_famgenus$genus, "-","_")
dnak_glmbin_gen.sigtab %>%
    left_join(select(rbind(dnak_glmbin_gen.acuta_dnak,dnak_glmbin_gen.oscillo_dnak),taxa,padj),by="taxa") %>%
    left_join(dnak_tax_famgenus,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
    mutate(taxa=str_replace_all(taxa, "g__",""))  %>%
    mutate(family=str_replace_all(family, "f__",""))  %>%
      select(-class)


```

## Significant genera for dnaK for binary model

Representation in boxes is according to the binary model


```{r, gyrB_bin_gen_plot,results='show',warning=FALSE,echo=FALSE,message=FALSE}

plot_grid(cag170_bin.boxplot, cag488_bin.boxplot, cag83_bin.boxplot, er4_bin.boxplot, ncol=2,nrow=2)

```

### gyrB

```{r, gyrB_bin_gen,results='show',warning=FALSE,echo=FALSE,message=FALSE}
gyrB_tax_famgenus <- as.data.frame(unique(phyloseq::tax_table(gyrB_physeq_dada2)[,c("class","family","genus")]))
gyrB_tax_famgenus$genus<- str_replace_all(gyrB_tax_famgenus$genus, "-","_")
gyrB_glmbin_gen.sigtab %>%
    left_join(gyrB_tax_famgenus,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
    mutate(taxa=str_replace_all(taxa, "g__",""))  %>%
    mutate(family=str_replace_all(family, "f__",""))  %>%
      select(-class)
```

### 16S

```{r, 16S_bin_gen_addfam,results='show',warning=FALSE,echo=FALSE,message=FALSE}
ssu_tax <- as.data.frame(unique(phyloseq::tax_table(ssu_physeq_dada2)[,c("class","family","genus")]))
ssu_tax$genus<- str_replace_all(ssu_tax$genus, "-","_")
ssu_glmbin_gen.sigtab %>%
    left_join(select(rbind(ssu_glmbin_gen.lachno_ssu,UBA1381_glmbin_gen.UBA1381_ssu,ssu_glmbin_gen.oscillo_ssu,cag138_glmbin_gen.cag138_ssu,cag74_glmbin_gen.cag74_ssu),taxa,padj),by="taxa") %>%
    left_join(ssu_tax,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
      select(-class)

```

## Continuous Model for Genera

### dnaK

```{r, dnak_cont_gen,results='show',warning=FALSE,echo=FALSE,message=FALSE}

dnak_glmcont_gen.sigtab %>%
    left_join(select(dnak_glmcont_gen.acuta_dnak,taxa,padj),by="taxa") %>%
    left_join(dnak_tax_famgenus,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
    mutate(taxa=str_replace_all(taxa, "g__",""))  %>%
    mutate(family=str_replace_all(family, "f__",""))  %>%
      select(-class)

```

### gyrB

```{r, gyrB_cont_gen,results='show',warning=FALSE,echo=FALSE,message=FALSE}
gyrB_glmcont_gen.sigtab %>%
    left_join(select(gyrB_glmcont_gen.genera_gyrB,taxa,p.adjust),by="taxa") %>%
    left_join(gyrB_tax_famgenus,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
    mutate(taxa=str_replace_all(taxa, "g__",""))  %>%
    mutate(family=str_replace_all(family, "f__",""))  %>%
      select(-class)

```

### 16S

```{r, 16S_cont_gen_addfam,results='show',warning=FALSE,echo=FALSE,message=FALSE}
ssu_glmcont_gen.sigtab %>%
    left_join(select(acuta_glmcont_gen.acuta_ssu,taxa,padj),by="taxa") %>%
    left_join(ssu_tax,by=c("taxa"="genus")) %>%
    filter(class%in%c("Clostridia","Clostridia_A","c__Clostridia"))  %>%
      select(-class)

```
